### 2021-09-30 GLOBI_bee_update_clean.Rmd
Adds script to pull bee / plant data from GLOBI. Also cleans data such that bees are source and plants are targets.

---
title:          Data Cleaning
output:         html_document
participants:   N. Bachelder, A. Chen, Z. Fang, M. Rapaport, S. Solomon
date:           Feb 18, 2021
source:         https://www.globalbioticinteractions.org/about
---

## Read data
```{r}
library(data.table)

df <- fread( 'https://zenodo.org/record/4460654/files/interactions.csv.gz')
```

# Filtering
```{r}
# List of total seven of bee families
b_families = c('Andrenidae','Apidae','Colletidae','Halictidae','Megachilidae','Melittidae','Stenotritidae')

# Bee as target, Plant as source
bee_source_df <- df %>% filter(sourceTaxonFamilyName %in% b_families, targetTaxonKingdomName == 'Plantae')

# Plant as source, Bee as target
plant_source_df = df %>% filter(sourceTaxonKingdomName == 'Plantae', targetTaxonFamilyName %in% b_families)

# Flip target/source for above df

colnames(plant_source_df) <- as.character(lapply(colnames(plant_source_df), textclean::swap, pattern1 = 'source', pattern2 = 'target'))

plant_source_df <- plant_source_df %>% rename(sourceCitation = targetCitation, sourceNamespace = targetNamespace, 
                                              sourceArchiveURI = targetArchiveURI, sourceDOI = targetDOI,
                                              sourceLastSeenAtUnixEpoch = targetLastSeenAtUnixEpoch)
```

## Combine
```{r}
# Combine without dropping duplicates
# The filtering method guarantees two disjoint data frames
interactions_dup = rbind(bee_source_df, plant_source_df)

# Combine and drop duplicates
interactions = rbind(bee_source_df, plant_source_df) %>% unique()
```

## Result
```{r}
interactions
```

# Export to csv.gz file
```{r}
write.csv(interactions, "generated_data/interactions.csv")
write.csv(interactions_dup, "generated_data/interactions_dup.csv")
```