### 2021-10-01 GLOBI-Fowler Data Prep.Rmd
Adds GLOBI-Fowler Modeling Prep. Also adds some GLOBI data exploration from last quarter (specifically on andrenidae bees)

---
title: "BeeModeling"
author: "Nick R. Bachelder"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)
```

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('interactions_dup.csv') %>% select(-X)

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}
```

```{r, echo = False}
beeCiteW
```

Define beeCites with varying criteria

```{r, fig.width=5,fig.height=2}
beeCiteNoFilter <- interactions_b %>% select(sourceTaxonSpeciesName, targetTaxonFamilyName) %>% dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% group_by(interestBee, interact) %>% count() %>% ungroup() %>% filter(interestBee != "", interact != "")

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(freq)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = freq, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(freq))

beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 5 citations to contribute to degree

beeIntDegreeW <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeW$numcitations <- beeCiteNoFilter$n
beeIntDegreeW$n_Bee <- beeCiteNoFilter$n_Bee
beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## weighted degree

beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type="inner") %>% filter(n > 5)

beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee", type="inner") %>% filter(n > 5)

beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee", type="inner") %>% filter(n > 5)
beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee", type="inner") %>% filter(n > 5)
```

## Include Fowler Data

```{r}
western_page <- read_html("https://jarrodfowler.com/pollen_specialist.html")
eastern_page <- read_html("https://jarrodfowler.com/specialist_bees.html")
central_page <- read_html("https://jarrodfowler.com/bees_pollen.html")

western_table <- western_page %>% html_table(fill=TRUE)
eastern_table<- eastern_page %>% html_table(fill=TRUE)
central_table<- central_page %>% html_table(fill=TRUE)

western_table <- as.data.frame(western_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
eastern_table <- as.data.frame(eastern_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
central_table <- as.data.frame(central_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)

western_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", western_table$Genus..Subgenus..species))
eastern_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", eastern_table$Genus..Subgenus..species))
central_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", central_table$Genus..Subgenus..species))

table <- rbind(western_table, eastern_table, central_table)
table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", table$Genus..Subgenus..species))

western_table$west_specialist <- 1
western_test <- western_table %>% dplyr::select(interestBee, west_specialist) %>% unique()

eastern_table$east_specialist <- 1
eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist) %>% unique()

central_table$cent_specialist <- 1
central_test <- central_table %>% dplyr::select(interestBee, cent_specialist) %>% unique()

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
table_test[is.na(table_test)] <- 0
```

```{r}
plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

fit = glm(specialist ~ n_Bee + degree, data = plotdf, family = 'binomial')

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 40))+ geom_jitter()+ scale_color_discrete(name = "Specialist") + geom_smooth()

write.csv(plotdf, 'BeeCiteRaw')
```
```{r}
plotdf <- merge(beeCiteC5, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist)), text = paste("Bee:", interestBee)) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity (Minimum 5 Citations) and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 15))+ geom_jitter() + scale_color_discrete(name = "Specialist")+ geom_smooth(se = FALSE)

write.csv(plotdf, 'BeeCiteC5')
```

```{r}
plotdf <- merge(beeCiteW, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Function of Degree')) + ggtitle(paste('Function of Degree and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 1)) + geom_smooth(se = FALSE)+ scale_color_discrete(name = "Specialist")

write.csv(plotdf, 'BeeCiteW')
```