### 2021-10-15 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
Refilters fowler so its not filtered only to GLOBI bees. Adds the pollen hosts as listed by GLOBI (to be compared to pollen hosts in MS data). Reruns tables to reflect changes.

---
title: "BeeModeling"
author: "Nick R. Bachelder"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)
```

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% select(-X)

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}
```

Define beeCites with varying criteria

```{r, fig.width=5,fig.height=2}
beeCiteNoFilter <- interactions_b %>% select(sourceTaxonSpeciesName, targetTaxonFamilyName) %>% dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% group_by(interestBee, interact) %>% count() %>% ungroup() %>% filter(interestBee != "", interact != "")

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(n)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(n))

beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 5 citations to contribute to degree

beeIntDegreeW <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeW$numcitations <- beeCiteNoFilter$n
beeIntDegreeW$n_Bee <- beeCiteNoFilter$n_Bee
beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## weighted degree

beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type = 'inner')

beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee",  type = 'inner')

beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee",  type = 'inner')

beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee",  type = 'inner')

beeCiteW
```

## Include Fowler Data

```{r}
western_page <- read_html("https://jarrodfowler.com/pollen_specialist.html")
eastern_page <- read_html("https://jarrodfowler.com/specialist_bees.html")
central_page <- read_html("https://jarrodfowler.com/bees_pollen.html")

western_table <- western_page %>% html_table(fill=TRUE)
eastern_table<- eastern_page %>% html_table(fill=TRUE)
central_table<- central_page %>% html_table(fill=TRUE)

western_table <- as.data.frame(western_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
eastern_table <- as.data.frame(eastern_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
central_table <- as.data.frame(central_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)

western_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", western_table$Genus..Subgenus..species))
eastern_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", eastern_table$Genus..Subgenus..species))
central_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", central_table$Genus..Subgenus..species))

table <- rbind(western_table, eastern_table, central_table)
table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", table$Genus..Subgenus..species))

western_table$west_specialist <- 1
western_test <- western_table %>% dplyr::select(interestBee, west_specialist) %>% unique()

eastern_table$east_specialist <- 1
eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist) %>% unique()

central_table$cent_specialist <- 1
central_test <- central_table %>% dplyr::select(interestBee, cent_specialist) %>% unique()

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
table_test[is.na(table_test)] <- 0

table_test$specialist <- table_test$west_specialist + table_test$east_specialist + table_test$cent_specialist
table_test$specialist <- as.integer(table_test$specialist > 0)
table_test$specialist[is.na(table_test$specialist)] <- 0

write.csv(table_test, 'fowler')
```

```{r}
plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

fit = glm(specialist ~ n_Bee + degree, data = plotdf, family = 'binomial')

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 40))+ geom_jitter()+ scale_color_discrete(name = "Specialist") + geom_smooth()

write.csv(plotdf, 'BeeCiteRaw')
```

```{r}
plotdf <- merge(beeCiteC5, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist)), text = paste("Bee:", interestBee)) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity (Minimum 5 Citations) and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 15))+ geom_jitter() + scale_color_discrete(name = "Specialist")+ geom_smooth(se = FALSE)

write.csv(plotdf, 'BeeCiteC5')
```

```{r}
plotdf <- merge(beeCiteW, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Function of Degree')) + ggtitle(paste('Function of Degree and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 1)) + geom_smooth(se = FALSE)+ scale_color_discrete(name = "Specialist")

write.csv(plotdf, 'BeeCiteW')
```