### 2022-01-26 MS_data_modeling/MS_data_explore.Rmd
Further work on paper

---
title: "MS_data_clean"
author: "Nick R. Bachelder"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lifecycle)
library(rvest)
library(dplyr)
library(tidyr)
library(plotly)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(kableExtra)
library(rcol)
library(plyr)
library(caret)

MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/UCSB Archive/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 2)

globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)

globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteC5') %>%
  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)

globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteW') %>%
  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)

fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/fowler') %>%
   mutate(interestBee = str_to_title(interestBee)) %>%
  dplyr::select(interestBee, west_specialist, east_specialist, cent_specialist, specialist) %>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))
```

```{r}
interactions <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv')

(interactions %>% filter(targetTaxonRank == 'family') %>% dplyr::select(sourceTaxonName, interactionTypeName, targetTaxonName) %>% 
    dplyr::rename('Bee Species' = sourceTaxonName, 'Interaction' = interactionTypeName, 'Plant Family' = targetTaxonName))[c(172, 112, 300, 124 ,5),]  %>%
  kbl() %>% kable_styling()

```

```{r}
MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
  dplyr::rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth', pollenHosts = 'Pollen hosts') %>%
  dplyr::select(interestBee, pollenHosts, degreeMS, specializationMS, References) %>%
  filter(!is.na(specializationMS)) %>% arrange(interestBee) %>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))

MS_data_format <- MS_data_format %>% filter(!(is.na(specializationMS)), !is.na(pollenHosts)) ### there

MS_data_format
```

### create family level column by pulling

```{r}
cp_nu_search(q='Sesamum', status = "accepted")$result$classification
```

```{r}
MS_data_format$pollenHosts <-  gsub("\\?.*", "", MS_data_format$pollenHosts)
MS_data_format$pollenHosts <-  gsub("\\..*", "", MS_data_format$pollenHosts)

ms_pollen_hosts = na.omit(unique(

  sub("\\/.*", "", word(unlist(str_split(unique(MS_data_format$pollenHosts), pattern = ', ')), 1))

  ))

## pulls all unique maximum level taxonomic names from MS pollen hosts
ms_pollen_hosts <- ms_pollen_hosts[!(ms_pollen_hosts == '')]
```

```{r}
library(rcol)
library(pbapply)

find_family <- function(name) {
  try <- try(cp_nu_search(q=name, status = "accepted"), silent=TRUE)
  if ("try-error" %in% class(try)) {
    label <- 'Unknown'
  }
  try <- try(str_to_title(tolower(
    (rbind( data.frame(try$result$classification[1]),
        data.frame(try$result$classification[2]),
       data.frame(try$result$classification[3])) %>% filter(rank == 'family'))[1,]$label
    )), silent=TRUE)
  if ("try-error" %in% class(try)) {
    label <- 'Unknown'
  }
  else {
    label <- try
  }
  return(label)
}

familys <- pblapply(ms_pollen_hosts, find_family)
```

```{r}
familys <- unlist(familys)
family_key <- data.frame(cbind(ms_pollen_hosts, familys)) %>% `colnames<-`(c('max_taxons', 'pollen_host_familys'))

family_key
```

## Make the below two entries have 0 rows with Avery's help
```{r}
family_key %>% filter(is.na(pollen_host_familys) | (pollen_host_familys == 'NA'))
```

```{r}
family_key %>% filter(pollen_host_familys == 'Unknown', grepl( ',', max_taxons, fixed = TRUE))
```

## make sure the below has no family name synonyms. Take out the unknownand NA values until we fix them with avery.

```{r}
pollen_family_MS <- MS_data_format %>% select(interestBee, pollenHosts) %>% separate_rows(pollenHosts, sep=', ') %>% mutate(max_taxons = word(pollenHosts, 1)  ) %>% left_join(family_key, by = 'max_taxons') %>% select(interestBee, pollen_host_familys) %>% 
  group_by(interestBee) %>% 
  dplyr::summarise(pollen_host_familys = paste(unique(as.vector(pollen_host_familys)), collapse = ', '))

MS_data_format <- MS_data_format %>% left_join(pollen_family_MS, by = 'interestBee') 
```

### Note, NAs are a result of there being no pollen hsots listed for a bee (should be fixed by filtering out bees with no pollen hosts), and Unknowns are a results of bees not being able to be found in catalougue of life

```{r}
avery_send <- MS_data_format %>% select(interestBee, pollenHosts, pollen_host_familys) %>% mutate(ifelse(pollen_host_familys == 'NA', NA, pollen_host_familys))

write.csv(avery_send, "/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Family_Level_Transform.csv")

avery_send
```

### Clean GLOBEE data

```{r}
pollenHostGLOBEE <- aggregate(interact~interestBee, globee_5cut, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
  dplyr::rename('pollenHostGLOBEE' = interact) %>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))
## this is plant interactions for each bee genus according to globi

globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE = degree) %>% 
  dplyr::select(interestBee, degreeGLOBEE, n_Bee) %>% unique() %>%
  arrange(interestBee) %>% merge(pollenHostGLOBEE, by = 'interestBee') %>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))

globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_C5 = degree) %>% 
  dplyr::select(interestBee, degreeGLOBEE_C5) %>% unique() %>%
  arrange(interestBee)%>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))

globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_T = degree) %>% 
  dplyr::select(interestBee, degreeGLOBEE_T) %>% unique() %>%
  arrange(interestBee)%>% 
  filter(grepl('^\\w+\\s\\w+$', interestBee))
```

```{r}
compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)  
compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), all = T) 
compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), all = T)

compare_df <- compare_df %>% merge(fowler %>% dplyr::rename(specialistFOWLER = specialist) %>% 
                                     dplyr::select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(specializationMS), 0, 1), 
         existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
         existsALL = existsGLOBEE*existsMS,
         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>%
  mutate(specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER))%>% 
  dplyr::rename(pollenHostsMS = pollenHosts ,pollenHostsFamilysMS = pollen_host_familys,ReferencesMS = References, nBeeGLOBEE = n_Bee) %>%
  dplyr::select(interestBee, specializationMS, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, specialistFOWLER, 
         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS, degreeMS) %>%
  rowwise() %>%
  dplyr::mutate(pollenHostsFamilysMS = ifelse(pollenHostsFamilysMS == 'NA', NA, pollenHostsFamilysMS) ) %>%
  dplyr::mutate(FamilyDegreeMS = ifelse(!is.na(pollenHostsFamilysMS), length(strsplit(pollenHostsFamilysMS, split = ', ')[[1]]), 0  ) )

compare_df %>% filter(existsALL == 1) %>% dplyr::select(interestBee, specializationMS, FamilyDegreeMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 'Degree MS' = FamilyDegreeMS, 'Raw Degree GLOBI' = degreeGLOBEE, 
                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE) %>%
  kbl() %>% kable_styling()
```

### Add individual citation cuts in addition to raw GLOBI
### Figure out what we want to define as 'specialist' MS data

```{r}
only_globi <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 0) %>% nrow()
globi_fowler <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 1) %>% nrow()
globi_MS <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 1, existsFOWLER == 0) %>% nrow()
MS_fowler <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 1) %>% nrow()
fowler_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 0, existsFOWLER == 1) %>% nrow()
MS_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 0) %>% nrow()
all_sources <- compare_df %>% filter(existsALL == 1) %>% nrow()

counts <- c(only_globi,fowler_only,MS_only, globi_MS, globi_fowler, MS_fowler, all_sources, nrow(compare_df))
data_sources <- c('Only in GLOBI', 'Only in Fowler', 'Only in MS', 'GLOBI and MS', 'GLOBI and Fowler', 'MS and Fowler', 'All Sources', 'Total')

data.frame(data_sources, counts) %>% mutate('Percent of All Bee Species Documented' = counts / nrow(compare_df)) %>% 
  dplyr::rename('Data Source' = data_sources, 'Unique Species Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
  kable_styling()
```

```{r}
compare_df %>% filter(existsALL == 1) %>% group_by(specializationMS) %>% dplyr::summarize('Mean GLOBI Degree'  = mean(degreeGLOBEE),
                                                                                   'GLOBI Degree IQR' = paste('[', round(quantile(degreeGLOBEE, 0.25),
                                                                                                                         2), 
                                                                                                                  ',', round(quantile(degreeGLOBEE, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Mean GLOBI C5 Degree' = mean(degreeGLOBEE_C5),
                                                                                   'GLOBI C5 Degree IQR' = paste('[', round(quantile(degreeGLOBEE_C5, 0.25), 2),
                                                                                                                  ',', round(quantile(degreeGLOBEE_C5, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Mean GLOBI T Degree' = mean(degreeGLOBEE_T),
                                                                                   'GLOBI T Degree IQR' = paste('[', round(quantile(degreeGLOBEE_T, 0.25), 2),
                                                                                                                  ',', round(quantile(degreeGLOBEE_T, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Percent Fowler Specialist' = mean(specialistFOWLER)
) %>% kbl(caption = 'Plant Family Degrees by MS Specialization') %>%
  kable_styling()
```

```{r}
# Bees fowler specialists but not MS specialists

odd_generalists <- compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5,degreeGLOBEE_T, degreeGLOBEE, pollenHostsFamilysMS, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 == 1)

odd_generalists

write.csv(odd_generalists, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/odd_polylectic.csv')
```

```{r}

(odd_generalists %>% dplyr:: select(interestBee, specializationMS, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 
                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE))[c(3,5),] %>% 
  kbl(caption = 'Odd Generalists') %>% kable_styling()

```

```{r}
# Bees NOT fowler specialists but MS specialists

odd_specialists <- compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, degreeGLOBEE_T, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 3)

write.csv(odd_specialists, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/odd_specialists.csv')
```

```{r}
(odd_specialists %>% dplyr:: select(interestBee, specializationMS, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 
                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE) )%>% 
  kbl(caption = 'Odd Specialists') %>% kable_styling()  
```

```{r}
# Bees NOT in MS

missing_MS <- compare_df %>% filter(existsMS == 0)

missing_MS

write.csv(missing_MS, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/missing_MS.csv')
```

## Plan: want to add pollen hosts from MS data so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column. ## DONE
NEED TO DISCUSS: The MS pollen hosts are a mix of genus and families, randomly. Maybe we can ask MS to seperate these into two columns, or just simply make a family level column only.

## Cleaning up source and pollen columns (removin author abbreviation)   ### DONE

## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa) ## WIP

## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables). ## DONE
DISCUSS: Turns out that we didnt cut globi bees, but we instead cut FOWLER bees. I unfiltered the fowler bees and it fixed the ratios. Now we can actually discuss data proportions

### Also: Added the target pollen hosts from globi so we can compare to pollen hosts from MS
This way we can directly compare the MS pollen hosts to the GLOBI pollen hosts. THe only issue is the above problem with a mix of taxonomic levels with the MS data

---------------

For next week:

Check out the web scraper to fix MS names to family level

Pull out data that is only in GLOBI and/or Fowler, use as group to check for old names as reason.

--------------

## Modeling below

# (1) Model same as old, except use conclusions from MS not fowler

```{r}
modeling_df <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% mutate(specialist = ifelse(specializationMS == 'Polylectic', 1, 0 )) %>% dplyr::select(interestBee, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, FamilyDegreeMS, specialist, pollenHostsFamilysMS, pollenHostGLOBEE)

modeling_df
ggplot(data= modeling_df, aes(x = nBeeGLOBEE, y = degreeGLOBEE_T, color = as.factor(specialist))) + geom_point() + xlim(0, 1000)
```
```{r}
length(unique(word(modeling_df$interestBee, 1)))

pollen_hosts_represented <- c(modeling_df$pollenHostsFamilysMS, modeling_df$pollenHostGLOBEE)
length(unique(pollen_hosts_represented))
```

```{r}
set.seed(123)
```

```{r}
calc_error_rate = function(predicted.value, true.value){
  return(mean(true.value!=predicted.value))
}
records = matrix(NA, nrow=3, ncol=2)
colnames(records) = c("train.error","test.error")
rownames(records) = c("tree","logistic","lasso")

nfold <- 10

n <- nrow(modeling_df)

in.trn <- sample.int(n, 0.8*n) 
trn.cl <- modeling_df[ in.trn,]
tst.cl <- modeling_df[-in.trn,]

folds <- sample(cut(1:nrow(trn.cl), breaks=nfold, labels=FALSE))
```

```{r}
library(tree)
library(maptree)

tree <- tree(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl)
cv <- cv.tree(tree, rand = folds, FUN = prune.misclass, K = 10)

best.size.cv = min(cv$size[cv$dev == min(cv$dev)])

tree.pruned <- prune.misclass(tree, best= best.size.cv)

train_pred_tree <- predict(tree.pruned, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
test_pred_tree <- predict(tree.pruned, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')

train_err <- calc_error_rate(train_pred_tree, trn.cl$specialist)
test_err <- calc_error_rate(test_pred_tree, tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
glmfit <- cv.glmnet(modeling_df %>% dplyr::select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), as.factor(modeling_df$specialist), family = 'binomial', alpha=1, 
          nfolds = 10)

coef  <-  coef(glmfit) ## uses lowest mse lambda
coef

predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")

predicttrain_log = ifelse(predicttrain_log > 0.5, 1, 0)
predicttest_log = ifelse(predicttest_log > 0.5, 1, 0)

trainerror <- calc_error_rate(predicttrain_log, trn.cl$specialist)
valerror <- calc_error_rate(predicttest_log, tst.cl$specialist)

1 - trainerror
1 - valerror
```

```{r}
library(randomForest)

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
library(randomForest)

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
library(randomForest)

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

# Try multi feature with best model:

```{r}
rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + degreeGLOBEE_T + degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err

misclassified <- tst.cl[tst.cl$specialist == predict_test,]
misclassified_vec <- tst.cl$specialist != predict_test

paste('Percent of misclassified that were specialists:', round(sum(misclassified$specialist) / nrow(misclassified), 3))
paste('Percent of studied that were specialists:', round(sum(modeling_df$specialist) / nrow(modeling_df), 3))
```

```{r}
tst.cl$misclass <- as.numeric(misclassified_vec)
splits <- c(0, 5, 10, 100, 100000)
tst.cl$nBeeBin <- as.factor(cut(tst.cl$nBeeGLOBEE, breaks=splits))
tst.cl %>% dplyr::select(misclass, nBeeBin) %>% group_by(nBeeBin) %>% dplyr::summarize(freq = n(), misclass_rate = mean(misclass))
```

# Try fitting to family degree:

```{r}
ggplot(data = modeling_df, aes(x= FamilyDegreeMS)) + geom_histogram()

splits <- c(0, 1, 2, 5, 10, 100)
modeling_df$MSFamilyDegreeBins <- cut(modeling_df$FamilyDegreeMS, breaks=splits)
modeling_df
```

```{r}
calc_error_rate = function(predicted.value, true.value){
  return(mean(true.value!=predicted.value))
}
records = matrix(NA, nrow=3, ncol=2)
colnames(records) = c("train.error","test.error")
rownames(records) = c("tree","logistic","lasso")

nfold <- 10

n <- nrow(modeling_df)

in.trn <- sample.int(n, 0.8*n) 
trn.cl <- modeling_df[ in.trn,]
tst.cl <- modeling_df[-in.trn,]

folds <- sample(cut(1:nrow(trn.cl), breaks=nfold, labels=FALSE))
```

```{r}
library(tree)
library(maptree)

tree <- tree(as.factor(MSFamilyDegreeBins) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl)
cv <- cv.tree(tree, rand = folds, FUN = prune.misclass, K = 10)

best.size.cv = min(cv$size[cv$dev == min(cv$dev)])

tree.pruned <- prune.misclass(tree, best= best.size.cv)

train_pred_tree <- predict(tree, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
test_pred_tree <- predict(tree, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')

train_err <- calc_error_rate(train_pred_tree, trn.cl$specialist)
test_err <- calc_error_rate(test_pred_tree, tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
glmfit <- cv.glmnet(modeling_df %>% dplyr::select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), as.factor(modeling_df$specialist), family = 'binomial', alpha=1, 
          nfolds = 10)

coef  <-  coef(glmfit) ## uses lowest mse lambda
coef

predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")

predicttrain_log = ifelse(predicttrain_log > 0.5, 1, 0)
predicttest_log = ifelse(predicttest_log > 0.5, 1, 0)

trainerror <- calc_error_rate(predicttrain_log, trn.cl$specialist)
valerror <- calc_error_rate(predicttest_log, tst.cl$specialist)

1 - trainerror
1 - valerror
```

```{r}
library(randomForest)

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
library(randomForest)

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
library(randomForest)

rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

# Try multi feature with best model:

```{r}
rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + degreeGLOBEE_T + degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)

predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")

train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)

test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)

1 - train_err
1 - test_err
```

```{r}
max(modeling_df$degreeGLOBEE_C5)
```

## Step 1: Classify bees as weird by finding misclassified bees by ML algoritem
## Step 2: Pull family level from both data and literature
## Step 3: Filter differences in family citations
## Step 4: Find similarities of family discrepencies (this identifies bias in data)
## Step 5: