### 2022-07-27 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
update plant names using the plant list

---
title: "BeeModeling"
author: "Nick R. Bachelder"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/colleen/Dropbox/Fall_2014/ucsb/Bee-Specialization-Modeling/")
knitr::opts_knit$set(error=T)
select = dplyr::select

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(tidyverse)
library(data.table)
library(readxl)
library(Taxonstand)

```

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('modeling_data/interactions_dup.csv')
ms <- read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 3) %>%
  rename_all(tolower) %>%
  mutate(scientificName = paste0(genus, ' ',sub('.*\\. ','',species))) %>%
  select(scientificName,everything()) %>%
  rename(diet_breadth=`redefined pollen diet breadth`)
fowler <- read_csv("modeling_data/fowler_hostplants.csv")

#who are the specialsits?
#let's say specialist the MS list and everything on fowler list that's missing from ms list
ms_oligo = ms$scientificName[ms$diet_breadth %in% c('Oligolectic','Monolectic')]
add_from_fowler = unique(fowler$scientificName[!fowler$scientificName %in% ms$scientificName])
specialists = c(ms_oligo,add_from_fowler)

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

# for (c in colnames(interactions_b)) {
#   interactions_b[[c]] <-  tolower(interactions_b[[c]])
# }
```

```{r}
#how many species in the missouri state data are in the globi data?
head(interactions_b)
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms$scientificName)]) #800 species in the ms data

#how many are specialists
ms_specialists = ms %>% filter(diet_breadth != 'Polylectic')
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_specialists$scientificName)]) #503 specialist species

#how many are generalists
ms_generalists = ms %>% filter(diet_breadth == 'Polylectic')
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_generalists$scientificName)]) #296 generalist species

```

```{r}
interactions_b[1:5,1:5]
```

Let's reformat and update the plant names in the interaction data 
```{r}
capitalize = function(a_string){

  beginning = toupper(substr(a_string,1,1))
  ending = substr(a_string,2,nchar(a_string))

  return(paste0(beginning,ending))

}

interactions_reformat = interactions_b %>% 
  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
         bee_family = capitalize(sourceTaxonFamilyName),
         plant_family = capitalize(targetTaxonFamilyName), 
         plant_genus = capitalize(targetTaxonGenusName),
         plant_species = capitalize(targetTaxonName)) %>%
  filter(scientificName != '' & plant_family !='' & plant_genus != "")

#

plant_df = interactions_reformat %>%
  distinct(plant_species)
test_me = plant_df$plant_species

# check_names = TPL(test_me) #takes forever to run, download output as csv
# write_csv(check_names,'modeling_data/plant_list_name_update.csv')
check_names = read_csv('modeling_data/plant_list_name_update.csv')

check_names %>% filter(Taxonomic.status != 'Accepted') %>% 
  select(Taxon,New.Genus,New.Species)

tpl_formatted = check_names %>% rename(plant_species = Taxon) %>% 
  mutate(accepted_name = ifelse(!is.na(New.Species),paste(New.Genus,New.Species),New.Genus) ) %>%
  mutate(accepted_name = ifelse(is.na(accepted_name), plant_species, accepted_name)) %>%
  select(plant_species,accepted_name)

tpl_formatted %>% filter(is.na(accepted_name))

#add accepted_name column to plant_df
interactions_updated = interactions_reformat %>% 
  group_by(scientificName,bee_family,plant_species,plant_family) %>%
  summarize(n=n()) %>%
  ungroup %>%
  left_join(tpl_formatted) %>%
  select(-plant_species) %>%
  rename(plant_species = accepted_name) %>%
  mutate(plant_genus = sub(" .*","",plant_species))

```

Let's do some reformatting
```{r}

globi_summ = interactions_updated %>%
  group_by(scientificName, plant_genus, plant_family,bee_family) %>%
  summarize(n_genus = sum(n)) %>% ungroup()

```

how many specialist bees do we have if we filter out interactions with <5 observations?
```{r}
globi_5n = globi_summ %>% filter(n_genus >= 5)
n_distinct(globi_5n$scientificName[globi_5n$scientificName %in% ms_specialists$scientificName]) #only 360

#how many other species
n_distinct(globi_5n$scientificName[!globi_5n$scientificName %in% ms_specialists$scientificName]) # 730, so about 1000 total

```
Calculate number of interaction partners

```{r}
#calculate the number of plant families each bee species is observed visiting
total_obs = globi_summ %>% group_by(scientificName) %>% summarize(n=sum(n_genus))

# degree_by_fam 
degree_by_fam = globi_summ %>% group_by(scientificName,plant_family) %>%
  summarize(n_family = sum(n_genus)) %>%
  left_join(total_obs) %>%
  mutate(freq_squared = (n_family/n)^2) %>% group_by(scientificName,n) %>%
  summarize(degree_family = n_distinct(plant_family), 
            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared))# %>%

#calculate the number of plant genera each bee species is observed visiting
degree_by_genus = globi_summ %>%
  left_join(total_obs) %>%
  mutate(freq_squared = (n_genus/n)^2) %>% 
  group_by(scientificName,bee_family) %>%
  summarize(degree_genus = n_distinct(plant_genus), 
            degree5_genus = sum(n_genus >= 5), weighted_degree_genus = sum(freq_squared))

globi_degree = degree_by_fam %>% left_join(degree_by_genus) %>%
  mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))
getwd()
# write_csv(globi_degree,'modelign_data/globi_degree.csv')

```