### 2023-12-21 modeling_data/old_files/Full Bredth v Specialization Study.Rmd
moving files

---
title: "Full Bredth v Specialization Study"
author: "Nick R. Bachelder"
date: "4/15/2022"
output: html_document
---

## issue fixes:

# Assume non-rare bees are 'not -rare' DONE
# Remove cases where bee rarity changes across area # DONE
# Change common catagory to 'not-rare' (assumption: fowler is exhaustive) # DONE
# location can be tossed due to data limitation, use broad scale east, west, central together.
# Do it at the genus level for plants

### aceae for families if we clean out family names for plants
### use tp for source

### Question: For and common rare bees in the easter, western, and central united states respectively, is the specialization indicative of limited diet bredth?
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
compare_df <- read.csv('modeling_data/compare_df.csv')

compare_df %>% filter(degreeMS > degreeMSGenus)
```

### lets experiment with all bees at FAMILY LEVEL

```{r}
monolectic <- compare_df %>% filter(specializationMS %in% c('Monolectic', 'Oligolectic'))
polylectic <- compare_df %>% filter(specializationMS == 'Polylectic')

monolectic
polylectic
```

### Note: correlation for polylectic

```{r}
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE, method=c("pearson"))
```

```{r}
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE_C5, method=c("pearson"))
```

## do at genus level

```{r}

```

```{r}

```

```
```

### lets experiment with all bees at GENUS LEVEL

```{r}
compare_df
```

### Note: monolectic all have degreeMS = 1, so cant measure correlation. polylectic all similar

```{r}
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE, method=c("pearson"))
```

```{r}
cor.test(monolectic$degreeMS, monolectic$degreeGLOBEE_C5, method=c("pearson"))
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE_C5, method=c("pearson"))
```

### Note: No correlation between MSdegree and GLOBEE_T degree for  monolectic, but one for polylectic

```{r}
cor.test(monolectic$degreeMS, monolectic$degreeGLOBEE_T, method=c("pearson"))
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE_T, method=c("pearson"))
```

```{r}
monolectic %>% ggplot(aes(x = degreeMS, y = degreeGLOBEE_T)) + geom_point()
```

```{r}
polylectic %>% ggplot(aes(x = degreeMS, y = degreeGLOBEE_T)) + geom_point()
```