### 2024-02-21 modeling_data/GLOBI_bee_update_clean.Rmd
adding script that got deleted

---
title:          Data Cleaning
output:         html_document
participants:   N. Bachelder, A. Chen, Z. Fang, M. Rapaport, S. Solomon, C. Smith
date:           Feb 18, 2021
source:         https://www.globalbioticinteractions.org/about
---

## Read data
```{r}
#library(data.table)
#df <- fread( 'https://zenodo.org/record/4460654/files/interactions.csv.gz')
library(vroom)
library(tidyverse)

df = vroom("all_bee_data.tsv")

```

# Filtering
```{r}
# List of total seven of bee families
b_families = c('Andrenidae','Apidae','Colletidae','Halictidae','Megachilidae','Melittidae','Stenotritidae')

# Bee as target, Plant as source
bee_source_df <- df %>% filter(sourceTaxonFamilyName %in% b_families, targetTaxonKingdomName == 'Plantae')

# Plant as source, Bee as target
plant_source_df = df %>% filter(sourceTaxonKingdomName == 'Plantae', targetTaxonFamilyName %in% b_families)

# Flip target/source for above df
colnames(plant_source_df) <- as.character(lapply(colnames(plant_source_df), textclean::swap, pattern1 = 'source', pattern2 = 'target'))
plant_source_df <- plant_source_df %>% dplyr::rename(sourceCitation = targetCitation, sourceNamespace = targetNamespace, 
                                              sourceArchiveURI = targetArchiveURI, sourceDOI = targetDOI,
                                              sourceLastSeenAtUnixEpoch = targetLastSeenAtUnixEpoch)
```

## Combine
```{r}
# Combine without dropping duplicates
# The filtering method guarantees two disjoint data frames
interactions_dup = rbind(bee_source_df, plant_source_df) 
# Combine and drop duplicates
interactions_b = rbind(bee_source_df, plant_source_df) %>% unique()
```

## Reformat column names
```{r}
for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

capitalize = function(a_string){

  beginning = toupper(substr(a_string,1,1))
  ending = substr(a_string,2,nchar(a_string))

  return(paste0(beginning,ending))

}

```

## Remove plants not id'd to at least genus level, and bees that are not id'd to species
## Change old family names for Asteraceae, Lamiaceae and Apiaceae
```{r}
#some of these are not vascular plants or not id'd to genus (e.g Heliantheae)
rm_me = c("Coleoptera","White","Picconiella","Heliantheae","Eupatorieae","Eremocarpus") #Eremocarpus is now split into 3 different genera

#reformat the interactions data:
interactions_reformat = interactions_b %>% 
  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
         bee_family = capitalize(sourceTaxonFamilyName),
         plant_family = capitalize(targetTaxonFamilyName), 
         plant_genus = capitalize(targetTaxonGenusName),
         plant_species = capitalize(targetTaxonName)) %>%
  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>%
  filter(targetTaxonRank %in% c('subspecies','species',"genus",'subgenus','variety')) %>%
  filter(sourceTaxonRank %in% c("subspecies",'variety','species')) %>%
  mutate(plant_family = ifelse(plant_family=='Compositae','Asteraceae',plant_family)) %>%
  mutate(plant_family = ifelse(plant_family=='Umbelliferae','Apiaceae',plant_family)) %>%
  mutate(plant_family = ifelse(plant_family=='Labiatae','Lamiaceae',plant_family)) %>% filter(!plant_species %in% rm_me)

```

# Export to csv file
```{r}
# readr::write_csv(interactions_reformat, "interactions-14dec2022.csv")
```